Unemployment Effects

### 2020-05-07 analysis.Rmd
added captions to my tables plus some other paper edits

library(rsconnect)
library(stargazer)
library(plyr)
library(plm)
library(wfe)
library(berryFunctions)
library(tidyverse)
```

@@

stargazer(shp.fit.pooling.tig.ext, shp.fit.pooling.control.tig.ext, shp.fit.pooling.swd.ext, shp.fit.pooling.control.swd.ext, shp.fit.pooling.ls.ext, shp.fit.pooling.control.ls.ext, liss.fit.pooling.tig.ext, liss.fit.pooling.control.tig.ext, liss.fit.pooling.swd.ext, liss.fit.pooling.control.swd.ext, liss.fit.pooling.ls.ext, liss.fit.pooling.control.ls.ext,
          type = "latex",
          out="./extension-table-1.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: pooled models with lagged unemployment",
          column.labels = c("(M1)", "(M2)", "(M3)", "(M4)", "(M5)", "(M6)", "(M7)", "(M8)", "(M9)", "(M10)", "(M11)", "(M12)"), 
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          dep.var.caption = c("SHP, LISS"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5)
          )
```

@@

stargazer(shp.fit.fe.tig.ext, shp.fit.fe.control.tig.ext, shp.fit.fe.swd.ext, shp.fit.fe.control.swd.ext, shp.fit.fe.ls.ext, shp.fit.fe.control.ls.ext, liss.fit.fe.tig.ext, liss.fit.fe.control.tig.ext, liss.fit.fe.swd.ext, liss.fit.fe.control.swd.ext, liss.fit.fe.ls.ext, liss.fit.fe.control.ls.ext,
          type = "latex",
          out="./extension-table-2.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          label = "table-2",
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: fixed effects models with lagged unemployment",
          column.labels = c("(M13)", "(M14)", "(M15)", "(M16)", "(M17)", "(M18)", "(M19)", "(M20)", "(M21)", "(M22)", "(M23)", "(M24)"),
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5)
          )
```

@@

  stargazer(p1, p1, p2, p2, p3, p3, p1, p1, p2, p2, p3, p3,
    order = c(1,2,6,3,4,5),
    coef = list(wfe_regression_data$c1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,1]),
                wfe_regression_data$c2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,1]),
                wfe_regression_data$c3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,1]),

@@

                wfe_regression_data$p4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,4]),
                wfe_regression_data$p5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,4]),
                wfe_regression_data$p6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,4])
             ),  
    type = "latex",
    out="./extension-table-3.tex",
    omit.stat = "all",
    notes = "SEs in parentheses. Columns 25-30 are from the SHP survey and columns 31-36 are from the LISS survey.",
    notes.append = TRUE,
    font.size = "tiny",
    digits = 2,
    column.sep.width = "-10pt", 
    title = "SHP Switzerland and LISS Netherlands: weighted fixed effects models",
    digit.separator="",
    rownames = FALSE,
    column.labels = c("(M25)", "(M26)", "(M27)", "(M28)", "(M29)", "(M30)", "(M31)", "(M32)", "(M33)", "(M34)", "(M35)", "(M36)"), 
    model.names = FALSE,
    model.numbers = FALSE,
    star.cutoffs=c(0.05,0.01,0.001),
    covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
    dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
    header = FALSE
    )  
```

>>>>

library(rsconnect)
library(stargazer)
library(plyr)
library(plm)
library(wfe)
library(berryFunctions)
library(kableExtra)
library(tidyverse)
```

@@

stargazer(shp.fit.pooling.tig.ext, shp.fit.pooling.control.tig.ext, shp.fit.pooling.swd.ext, shp.fit.pooling.control.swd.ext, shp.fit.pooling.ls.ext, shp.fit.pooling.control.ls.ext, liss.fit.pooling.tig.ext, liss.fit.pooling.control.tig.ext, liss.fit.pooling.swd.ext, liss.fit.pooling.control.swd.ext, liss.fit.pooling.ls.ext, liss.fit.pooling.control.ls.ext,
          type = "latex",
          table.placement = "H",
          out="./extension-table-1.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: pooled models with lagged unemployment",
          column.labels = c("(M1)", "(M2)", "(M3)", "(M4)", "(M5)", "(M6)", "(M7)", "(M8)", "(M9)", "(M10)", "(M11)", "(M12)"), 
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5),
          notes = "\\parbox[t]{15cm}{The above table presents the results from a series of pooled effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 1-6 are regression outputs from the SHP survey data while columns 7-12 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficients are statistically across all questions for both surveys with and without controls. However, the lagged unemployment coefficient is only statistically significant for the SHP data and the interaction coefficient is only significant for the life satisfaction question, for both surveys. }"
          )
```

@@

stargazer(shp.fit.fe.tig.ext, shp.fit.fe.control.tig.ext, shp.fit.fe.swd.ext, shp.fit.fe.control.swd.ext, shp.fit.fe.ls.ext, shp.fit.fe.control.ls.ext, liss.fit.fe.tig.ext, liss.fit.fe.control.tig.ext, liss.fit.fe.swd.ext, liss.fit.fe.control.swd.ext, liss.fit.fe.ls.ext, liss.fit.fe.control.ls.ext,
          type = "latex",
          out="./extension-table-2.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          label = "table-2",
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: fixed effects models with lagged unemployment",
          column.labels = c("(M13)", "(M14)", "(M15)", "(M16)", "(M17)", "(M18)", "(M19)", "(M20)", "(M21)", "(M22)", "(M23)", "(M24)"),
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5),
          notes = "\\parbox[t]{15cm}{The above table presents the results from a series of fixed effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 13-18 are regression outputs from the SHP survey data while columns 19-24 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficients are statistically significant for the life satisfaction question for both surveys with and without controls. However, the lagged unemployment coefficient is only statistically significant for the satisfaction with democracy question on the SHP survey while the interaction coefficient is significant for the satisfaction with democracy on the LISS survey.}"
          )
```

@@

  stargazer(p1, p1, p2, p2, p3, p3, p1, p1, p2, p2, p3, p3,
    coef = list(wfe_regression_data$c1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,1]),
                wfe_regression_data$c2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,1]),
                wfe_regression_data$c3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,1]),

@@

                wfe_regression_data$p4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,4]),
                wfe_regression_data$p5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,4]),
                wfe_regression_data$p6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,4])
             ),  
    type = "latex",
    out="./extension-table-3.tex",
    omit.stat = "all",
    notes = "\\parbox[t]{15cm}{The above table presents the results from a series of weighted pooled effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 25-30 are regression outputs from the SHP survey data while columns 31-36 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficient is statistically significant for the life satisfaction question on both surveys with and without controls. However, the lagged unemployment coefficient and the interaction coefficient are not statistically significant for any of the questions on either survey.}",
    font.size = "tiny",
    digits = 2,
    column.sep.width = "-10pt", 
    title = "SHP Switzerland and LISS Netherlands: weighted fixed effects models",
    digit.separator="",
    rownames = FALSE,
    column.labels = c("(M25)", "(M26)", "(M27)", "(M28)", "(M29)", "(M30)", "(M31)", "(M32)", "(M33)", "(M34)", "(M35)", "(M36)"), 
    model.names = FALSE,
    model.numbers = FALSE,
    star.cutoffs=c(0.05,0.01,0.001),
    covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
    dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
    header = FALSE,
    order = c(1,2,6,3,4,5)
    )  
```

### 2020-05-07 paper.Rmd
added captions to my tables plus some other paper edits

library(rsconnect)
library(stargazer)
library(plyr)
library(plm)
library(wfe)
library(tidyverse)
```

@@

Bauer's work draws on the theoretical concepts of James Laurence who argued that the loss of one's job may lead to a deterioration of their trust in the economic system (@laurence2015). Furthermore, the loss of one's job brings people into closer connect with bureaucratic institutions and government social programs. Such interactions may incite a frusturation with the way in which the government supports individuals experiencing joblessness and could lead to a dissolusionment with the current government (@waters2007). It is from this theoretical basis that Bauer hypothesizes unemployment to be negatively associated with trust in government and satisfaction with democracy. There is, however, an alternative theory. Should a person feel supported and appreciated by the government and its programs during their period of joblessness, their feelings towards the government may not change or may even improve (@roth2009).

@@

The fundamental motivation beyond Bauer’s paper is to study how unemployment affects trust in government and satisfaction with democracy. The goal of my extension to provide further insights and clarity into that relationship.

@@

The first regression, displayed in Table 1, illustrates the results of the pooled model. In that model, unemployed is statistically significant across all questions of both surveys. Lagged unemployed is statistically significant across all questions from the SHP survey but is only statistically significant for the life satisfaction question of the LISS survey. The interaction term is only statistically significant for the life statisfaction question of each survey.

@@

Because the data provided does not provide the ability to control for all observable and unobservable confounding variables arising from time and household variation, the second regression, displayed in Table 2, makes use of a fixed effects model. There are strong differences between the pooled model and the fixed effects model. The size of the coefficients has dropped across all of the covariates by a considerable amount. As such, fewer coefficients are statistically significant. Unemployed is no longer statistically significant, aside from the life satisfaction questions. Lagged unemployed remains statistically significant for satisfaction with democracy while the interaction term is only statistically significant for life satisfaction.

@@

Finally, I conduct the regression analysis using the weighted fixed effects model used by Bauer. The coefficients have remained similar, in size, to the fixed effects model. However, lagged unemployed and the interaction between unemployed and lagged unemployed are no longer statistically significant for any question. Unemployed is statistically significant for the life satisfaction question.

@@

My findings largely confirmed Bauer's. Unemployment does not appear to have a statistically significant effect on trust in government or satisfaction with democracy. The interaction term between unemployment in the present year and unemployment in the previous year was not statistically significant. The lack of statistical significance indicates that the aggregation of two years of unemployment does not create significant impacts on one's views towards government and democracy. However, in the dataset from Switzerland (SHP) there was a statistically significant negative relationship between satisfaction with democracy and lagged unemployment. This relationship is only present in the pooled model and the fixed effects model. Because the relationship between lagged unemployment and satisfaction with democracy is not present in the weighted fixed effects model, supposedly the most robust model, it is unclear how strong that finding is.

@@

In 2018, Paul Bauer published the paper, "Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation" in which he explored the relationship between a person's unemployment status and their trust in government and satisfaction with democracy (@bauer2018). Using data from the SHP and LISS household surveys, Bauer found that there was no statistically significant effect between unemployment and either trust in government or statisfaction with democracy. These findings challenged his hypothesis that short-term unemployment would people to loss faith in the government and its institutions. Bauer's hypothesis was based on theory and prior findings of a link between unemployment and decreasing life satisfcation.

Using RStudio, I took Bauer's data and models and attempted to explore the possibility of heterogenous effects between short-term and long-term unemployment. I took advantage of the panel data to create a variable called lagged unemployment, which indicated whether a household had been unemployed in the previous year. I included the lagged unemployment variable in Bauer's regression models as an interaction term with unemployment. I was very interested in the coefficient on the interaction term (signifying a household that was unemployed in both the previous year and the current year). I assumed, like Bauer, that the extended period of unemployment would lead to a decrease in their trust in government and satisfaction with democracy. The results from the regressions disproved that hypothesis and reaffirmed Bauer's findings. Again, there were no statistically significant effects from the longer period of unemployment on either trust in government or satisfaction with democracy.

Both my results, and Bauer's, can offer encouraging interpretations. As Felix Roth theorized, positive interactions with government programs and institutions arising from one's unemployment may in fact lead to improvements in that individuals feelings towards the government, or at the very least no real change (@roth2009). It is possible that the lack of statistically significant effects in the regression models is due to highly effective government programs for those experiencing joblessness in Switzerland and the Netherlands. However, it could also be due to limitations with the data. Regardless, futher studies making use of longer term data across several countries are warranted to more conclusively understand the link between unemployment and feelings towards government.

>>>>

library(rsconnect)
library(stargazer)
library(plyr)
library(plm)
library(wfe)
library(kableExtra)
library(tidyverse)
```

@@

Bauer's work draws on the theoretical concepts of James Laurence who argued that the loss of one's job may lead to a deterioration of their trust in the economic system (@laurence2015). Furthermore, the loss of one's job brings people into closer contact with bureaucratic institutions and government social programs. Such interactions may incite a frusturation with the way in which the government supports individuals experiencing joblessness and could lead to a dissolusionment with the current government (@waters2007). It is from this theoretical basis that Bauer hypothesizes unemployment to be negatively associated with trust in government and satisfaction with democracy. There is, however, an alternative theory. Should a person feel supported and appreciated by the government and its programs during their period of joblessness, their feelings towards the government may not change or may even improve (@roth2009).

@@

The fundamental motivation behind Bauer’s paper is to study how unemployment affects trust in government and satisfaction with democracy. The goal of my extension to provide further insights and clarity into that relationship.

@@

The first regression, displayed in Table 1, illustrates the results of the pooled model. In that model, 'Unemployed' is statistically significant across all questions from both surveys. There is a negative relationship unemployment and satisfaction with the government. "Lagged unemployed" is statistically significant across all questions from the SHP survey but is only statistically significant for the life satisfaction question of the LISS survey. The coefficients are negative but smaller in magnitude than they are for "Unemployed". The interaction term is only statistically significant for the life statisfaction question of each survey. Furthermore, the coefficients on the trust in government and satisfaction with democracy questions from the SHP survey are positive, though not significant.

@@

Because the data made available does not provide the ability to control for all observable and unobservable confounding variables arising from time and household variation, the second regression, displayed in Table 2, makes use of a fixed effects model. There are strong differences between the pooled model and the fixed effects model. The size of the coefficients has dropped across all of the covariates by a considerable amount. As such, fewer coefficients are statistically significant. 'Unemployed' is no longer statistically significant, aside from the life satisfaction questions. 'Lagged unemployed' remains statistically significant for satisfaction with democracy while the interaction term is only statistically significant for life satisfaction.

@@

Lastly, I conduct the regression analyses using the weighted fixed effects model used by Bauer. The coefficients have remained similar, in size, to the fixed effects model. However, 'Lagged unemployed' and the interaction between 'Unemployed' and 'Lagged unemployed' are no longer statistically significant for any question. 'Unemployed' is statistically significant for the life satisfaction question.

@@

My findings largely confirmed Bauer's. Unemployment does not appear to have a statistically significant effect on trust in government or satisfaction with democracy. The interaction term between unemployment in the present year and unemployment in the previous year was not statistically significant. However, in the dataset from Switzerland (SHP) there was a statistically significant negative relationship between satisfaction with democracy and lagged unemployment. This relationship is only present in the pooled model and the fixed effects model. Because the relationship between lagged unemployment and satisfaction with democracy is not present in the weighted fixed effects model, supposedly the most robust model, it is unclear how strong that finding is. Ultimately though, the lack of statistical significance for the interaction term indicates that the aggregation of two years of unemployment does not create significant impacts on one's views towards government and democracy.

@@

In 2018, Paul Bauer published the paper, "Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation" in which he explored the relationship between a person's unemployment status and their trust in government and satisfaction with democracy (@bauer2018). Using data from the SHP and LISS household surveys, Bauer found that there was no statistically significant effect between unemployment and either trust in government or statisfaction with democracy. These findings challenged his hypothesis that short-term unemployment would lead people to lose faith in the government and its institutions. Bauer's hypothesis was based on theory and prior findings indicating a link between unemployment and decreasing life satisfaction.

Using RStudio, I took Bauer's data and models and attempted to explore the possibility of heterogenous effects between short-term and long-term unemployment. I took advantage of the panel data to create a variable called lagged unemployment, which indicated whether a household had reported unemployment in the previous year. I included the lagged unemployment variable in Bauer's regression models as an interaction term with unemployment. I was very interested in the coefficient on the interaction term (signifying a household that was unemployed in both the previous year and the current year). I assumed, like Bauer, that the extended period of unemployment would lead to a decrease in a household's trust in government and satisfaction with democracy. The results from the regressions disproved that hypothesis and reaffirmed Bauer's findings. Again, there were no statistically significant effects from the longer period of unemployment on either trust in government or satisfaction with democracy.

Both my results, and Bauer's, can offer encouraging interpretations. As Felix Roth theorized, positive interactions with government programs and institutions during a period of unemployment may lead to improvements in that individual's feelings towards the government, or, at the very least, prevent a strong decrease in attitude (@roth2009). Thus, the lack of statistically significant effects in these regression models may indicate highly effective programs for those experiencing joblessness. Looking at the countries surveyed, Switzerland and the Netherlands, such an optimistic interpretation is plausible given the robustness of their social welfare programs. However, the lack of a consistent effect could also stem from limitations with the data. As such, futher studies making use of longer term data across several countries are warranted to more conclusively understand the link between unemployment and feelings towards government.
